---
title: "Meta Analysis"
author: "Pamela Inostroza"
date: "28/05/2020"
output: 
  html_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# if (!require("devtools")) {
#   install.packages("devtools")
# }
# devtools::install_github("MathiasHarrer/dmetar")

knitr::opts_chunk$set(echo = TRUE)
options(width = 10000)

library(dplyr)
library(meta)
library(metafor)
library(rgl)
library(netmeta)
library(dmetar)


se.from.p(effect.size = -4.2, 
          p = 0.02, 
          N = 11,
          effect.size.type= "difference")

dire <- "G:/My Drive/Master in Statistics/Meta analysis/Paper assignment Meta Analysis/code and drafts and tables/"
# 1. Read in the data
data <- read.csv(paste0(dire,"data.csv"), as.is=TRUE,header = TRUE)
data$dose <- paste(data$Doses.mg,"mg")
data$dosec <- ifelse(data$Doses.mg < 0.5, "Dose < 0.5 mg",ifelse(data$Doses.mg > 0.5, "Dose > 0.5 mg", "Dose = 0.5 mg"))


```

Data containig the following information:  

- the title,
- the author,
- the year of publication,
- Ne, the Number of patients (sample size) in the experimental treatment group,
- Me, the Mean response among the patients in the experimental group,
- SDe, the Standard deviation of the response in the experimental group, and
- Nc, Mc and SDc the sample size, mean response and standard deviation in the control patients.

Analysis was done following https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/

# Meta Analysis for all studies

```{r model1,results= 'markup'}

# #1. Calculate mean difference, variance and weights
# MD <- with(data, Me - Mc)
# varMD <- with(data, SDe^2/Ne + SDc^2/Nc)
# weight <- 1/varMD
# # 2. Calculate the inverse variance estimator
# round(weighted.mean(MD, weight), 4)
# 
# # 3. Calculate the variance
# round(1/sum(weight), 4)
# round(c(mc1$TE.fixed, mc1$seTE.fixed^2), 4)
# mc1$w.fixed[1]
# sum(mc1$w.fixed)
# round(100*mc1$w.fixed[1] / sum(mc1$w.fixed), 2)

# 1. Apply generic inverse variance method
# mc1.gen <- metagen(mc1$TE, mc1$seTE, sm="MD")
# print(mc1.gen, digits=2)

# # 3. Print results for fixed effect and random effects method
# c(mc1$TE.fixed, mc1$TE.random)
# c(mc1.gen$TE.fixed, mc1.gen$TE.random)

##########################################
#Fixed effect model
##########################################
#Mean difference 
mc1 <- metacont(Ne, Me, SDe, Nc, Mc, SDc, data=data, studlab=paste(Author,dose))
print(mc1, digits=2)
 
forest(mc1, comb.random=TRUE, xlab="Difference in mean response (intervention - control)", 
  leftcols = c("studlab"), digits.sd = 1,digits = 1,
       prediction = TRUE,
        plotwidth = "5cm", spacing = 1, fontsize = 9,
        colgap = "0.3mm",
        xlim=c(-20,10), xlab.pos=0, smlab.pos=0)


#Standardized mean difference Hedges
mc2 <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD",  data=data, studlab=paste(Author,dose), method.smd = "Hedges")
#Standardized mean difference Cohen's d
mc2 <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD",  data=data, studlab=paste(Author,dose), method.smd = "Cohen")
#round(c(mc2$TE.fixed, mc2$seTE.fixed^2), 4)
print(mc2, digits=2)

mc_author <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD",  data=data, studlab=paste(Author), method.smd = "Cohen")
#round(c(mc2$TE.fixed, mc2$seTE.fixed^2), 4)
print(mc_author, digits=2)


#Save Cohen's d and standard error
data$TE <- mc2$TE
data$seTE <- mc2$seTE

##########################################
#Random effect model
##########################################
mc2.MD <- metacont(Ne, Me, SDe, Nc, Mc, SDc, data=data, comb.fixed=FALSE, hakn=TRUE, prediction=TRUE, studlab=paste(Author,dose))
print(mc2.MD, digits=2)

mc2.sMD <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD", data=data, comb.fixed=FALSE, hakn=TRUE, prediction=TRUE, method.smd = "Cohen", studlab=paste(Author,dose))
print(mc2.sMD, digits=2)

forest(mc2.sMD, prediction=TRUE, col.predict="black",comb.random=TRUE, 
       xlab="Standardized mean difference, Cohen's D", 
       digits.sd = 1,digits = 1, sortvar = TE,
       plotwidth = "5cm", spacing = 1, fontsize = 9,
       colgap = "0.3mm",
       xlim=c(-5,5), xlab.pos=0, smlab.pos=0)

sgroup <- subgroup.analysis.mixed.effects(x = mc2.sMD,
                                subgroups = data$group)
forest(sgroup,digits.sd = 1,digits = 1,
       prediction = TRUE,
        plotwidth = "5cm", spacing = 0.8, fontsize = 9,
        colgap = "0.3mm",
       xlim=c(-5,5), xlab.pos=0, smlab.pos=0)

sdose <- subgroup.analysis.mixed.effects(x = mc2.sMD,
                                subgroups = data$dosec)
forest(sdose,digits.sd = 1,digits = 1,comb.random=TRUE,
       prediction = TRUE, 
        plotwidth = "5cm", spacing = .8, fontsize = 9,
        colgap = "0.3mm", xlim=c(-5,5),)

```


```{r graph, results='markup'}

trimfill(mc2.sMD,comb.random=TRUE, hakn =TRUE)
forest(trimfill(mc2.sMD,comb.random=TRUE),digits.sd = 1,digits = 1,digits.seTE = 1,
       prediction = TRUE,leftcols = c("studlab"),
       xlim=c(-5,5), 
       plotwidth = "5cm", spacing = .8, fontsize = 9,
       colgap = "5mm")
par(mfrow=c(1,2))
funnel(mc2.sMD)
title("Studies included")
funnel(trimfill(mc2.sMD,comb.random=TRUE)) #Trim-and-fill method for estimating and adjusting for the number and outcomes of missing studies
title("Trim-and-fill")

FEM <- rma(yi=TE, sei= seTE, data=data,method="FE")
summary(FEM)
REM <- rma(yi=TE, sei= seTE, data=data, method="REML")
summary(REM)

regtest(REM,model = 'lm', predictor = 'se')# Regresion test, Egger's intercept test
ranktest (REM) #Begg's test
fsn(TE, seTE, data =data) #failsafeN


```

# Diagnostic, outliers

```{r met2,results= 'markup'}

fo <- find.outliers(mc2)
fo

forest(fo, col.predict = "blue",
       leftcols = c("studlab"),digits.sd = 1,digits = 1,
       prediction = TRUE,
        plotwidth = "5cm", spacing = 1, fontsize = 9,
        colgap = "0.3mm")
inf.analysis <- InfluenceAnalysis(x = mc2, random = TRUE)
summary(inf.analysis)
plot(inf.analysis, "influence")
plot(inf.analysis, "baujat")
plot(inf.analysis, "es")


m.sMD <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD", data=data, hakn=TRUE, prediction=TRUE,
        method.tau = "SJ", comb.random = TRUE, method.smd = "Cohen", studlab=paste(Author,dose), exclude = c(3))
m.sMD
REM2 <- rma(yi=TE, sei= seTE, data=data, method="REML",subset = c(1,2,4,5,6,7,8,9,10))
summary(REM2)

# Influential studies?
leave1out(REM2)

# Checking normality assumption
qqnorm(REM)
qqnorm(REM2)

# getting empirical Bayes estimates for individual study effects
blup(REM)
```

# Regressions 

```{r reg,results= 'markup'}
# convergence over time?
x<-cumul(REM, order=order(data$Year)) #order according to publication date
forest(x,cex = 0.8)

# Fixed effects model with Doses as a categorical moderator (ANOVA-model)
FEMANOVA11<-rma(TE, seTE, data=data, mods=~factor(Year), method="FE")
summary(FEMANOVA11)
# Random effects model with Weeks as a categorical moderator (ANOVA-model)
REMANOVA11<- rma(TE, seTE, data=data, mods=~factor(Year), method="REML")
summary(REMANOVA11)
output.metareg1 <- metareg(mc2,Year, method="REML")
summary(output.metareg1)


FEMANOVA21<-rma(TE, seTE, data=data, mods=~factor(dose), method="FE")
summary(FEMANOVA21)
FEMANOVA22<-rma(TE, seTE, data=data, mods=~Doses.mg, method="FE")
summary(FEMANOVA22)
REMANOVA21<- rma(TE, seTE, data=data, mods=~factor(dose), method="REML")
summary(REMANOVA21)
REMANOVA22<- rma(TE, seTE, data=data, mods=~Doses.mg, method="REML")
summary(REMANOVA22)
output.metareg2 <-metareg(mc2,Doses.mg)
summary(output.metareg2)

output.metareg3 <- metareg(mc2,Female, method="REML")
summary(output.metareg3)


output.metareg4 <- metareg(mc2,Age, method="REML")
summary(output.metareg4)

par(mfrow=c(2,2))
bubble(output.metareg1,
       xlab = "Publication Year",
       col.line = "blue",xlim=c(2008,2020), ylim = c(-2,0), cex.studlab = 0.6,
       pos.studlab = 3, studlab = TRUE, ylab = "SMD")

bubble(output.metareg2,
       xlab = "Doses",
       col.line = "blue",cex.studlab = 0.6,xlim=c(0.1,0.9), ylim = c(-2,0), 
       pos.studlab = 3, studlab = TRUE, ylab = "SMD")

bubble(output.metareg3,
       xlab = "Female (%)",
       col.line = "blue",xlim=c(0.4,0.8), ylim = c(-2,0), cex.studlab = 0.6,
       pos.studlab = 3, studlab = TRUE, ylab = "SMD")

bubble(output.metareg4,
       xlab = "Age (mean)",
       col.line = "blue",xlim=c(42,46), ylim = c(-2,0), cex.studlab = 0.6,
       pos.studlab = 3, studlab = TRUE, ylab = "SMD")


data$Age_sc <- scale(data$Age, center=TRUE, scale=FALSE) 
data$Female_sc <- scale(data$Female, center=TRUE, scale=FALSE) 
data$Year_sc <- scale(data$Year, center=TRUE, scale=FALSE) 
data$Doses_sc <- scale(data$Doses.mg, center=TRUE, scale=FALSE) 

mc2 <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD",  data=data, studlab=paste(Author,dose), method.smd = "Cohen")
print(mc2, digits=2)

library(PerformanceAnalytics)
chart.Correlation(data[,c("TE","Year","Doses.mg","Female","Age")])


metareg1 <- rma(yi=TE,sei=seTE, data=data,mods = ~ Year_sc+Doses_sc+group+Female_sc * Age_sc, method="REML",test = "knha")
summary(metareg1)

metareg2 <- rma(yi=TE,sei=seTE, data=data,mods = ~ Year_sc+Doses_sc+Female_sc * Age_sc, method="REML",test = "knha")
summary(metareg2)

anova(metareg1,metareg2)

permutest(metareg2)

multimodel.inference(TE = "TE", 
                     seTE = "seTE",
                     data = data,
                     predictors = c("Year_sc","Doses_sc","Female_sc","Age_sc"),
                     interaction = TRUE)

```

```{r gos,results= 'hide'}
dat.gosh <- gosh(REM)

```

```{r plot,results= 'markup'}
plot(dat.gosh, alpha= 0.1, col = "blue")
gosh.diagnostics(dat.gosh)

```

# Multilevel model 

```{r multilevel,results= 'markup'}
full.model<-rma.mv(TE, seTE^2, data=data, random=~1|Author, 
                   tdist = TRUE, 
                   method = "REML") #note that we use rma.mv rather than rma 

summary(full.model) ## shows the results

multilevel<-rma.mv(TE, seTE^2, data=data, random=~1|Author, 
                   tdist = TRUE, 
                   mods = ~ Year_sc  + Female_sc * Age_sc,
                   method = "REML",sigma2 = c(0)) #note that we use rma.mv rather than rma 

summary(multilevel) ## shows the results

forest(full.model, order="obs")

```

# Meta Analysis one dose (around 0.5 mg ketamine)

```{r model2,results= 'markup'}

data1 <- data[!data$Doses.mg %in% c(0.28,0.84,0.20),]

#Mean difference 
mc1 <- metacont(Ne, Me, SDe, Nc, Mc, SDc, data=data1, studlab=paste(Author))
print(mc1, digits=2)
forest(mc1, comb.random=TRUE, xlab="Difference in mean response (intervention - control)",
        leftcols = c("studlab"),  digits.sd = 1,digits = 1,
       prediction = TRUE,
        plotwidth = "5cm", spacing = 1, fontsize = 9,
        colgap = "0.3mm", 
        xlim=c(-20,20), xlab.pos=0, smlab.pos=0)

#Standardized mean difference Cohen's d
mc2 <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD",  data=data1, studlab=paste(Author), method.smd = "Cohen")
print(mc2, digits=2)

data1$TE <- mc2$TE
data1$seTE <- mc2$seTE

##########################################
#Random effect model
##########################################
mc2.MD <- metacont(Ne, Me, SDe, Nc, Mc, SDc, data=data1, comb.fixed=FALSE, hakn=TRUE, prediction=TRUE, studlab=paste(Author))
print(mc2.MD, digits=2)

mc2.sMD <- metacont(Ne, Me, SDe, Nc, Mc, SDc, sm="SMD", data=data1, comb.fixed=FALSE, hakn=TRUE, prediction=TRUE, method.smd = "Cohen", studlab=paste(Author))
print(mc2.sMD, digits=2)

forest(mc2.sMD, prediction=TRUE, col.predict="black",comb.random=TRUE, xlab="Standardized mean difference, Cohen's D", 
        leftcols = c("studlab"),  digits.sd = 1,digits = 1,
        plotwidth = "5cm", spacing = 1, fontsize = 9,
        colgap = "0.3mm",
       xlim=c(-10,10), xlab.pos=0, smlab.pos=0)

funnel(mc2.sMD)
trimfill(mc2.sMD,comb.random=TRUE)
funnel(trimfill(mc2.sMD,comb.random=TRUE)) #Trim-and-fill method for estimating and adjusting for the number and outcomes of missing studies

FEM <- rma(yi=TE, sei= seTE, data=data1,method="FE")
summary(FEM)
REM <- rma(yi=TE, sei= seTE, data=data1, method="REML")
summary(REM)

regtest(REM,model = 'lm', predictor = 'se')# Regresion test, Egger's intercept test
ranktest (REM) #Begg's test
fsn(TE, seTE, data =data1) #failsafeN

output.metareg1 <- metareg(mc2,Year, method="REML")
summary(output.metareg1)

bubble(output.metareg1,
       xlab = "Publication Year",
       col.line = "blue",xlim=c(2008,2020), ylim = c(-4,-0.5), cex.studlab = 0.6,
       pos.studlab = 3, studlab = TRUE)
``` 

# Power analysis  

```{r power,results= 'markup'}

power.analysis(d=1.96, 
                k=2, 
                n1=15, 
                n2=15, 
                p=0.001,
               heterogeneity = "low") #"low", "moderate","high"
```

# Network

```{r network,results= 'markup'}


m.netmeta <- netmeta(TE = TE,
                     seTE = seTE,
                     treat1 = treat1,
                     treat2 = treat2,
                     studlab = paste(Author,dose),
                     data = data,
                     sm = "MD",
                     comb.fixed = TRUE,
                     comb.random = FALSE,
                     reference.group = "plac",
                     details.chkmultiarm = TRUE,
                     sep.trts = " vs ")
m.netmeta
decomp.design(m.netmeta)

netgraph(m.netmeta)

direct.evidence.plot(x = m.netmeta)
```
