---
title: "Structural equations modeling"
author: "Pamela Inostroza"
date: "10/05/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(psych)
library(expss)
library(sjlabelled)
library(dplyr)

library(semPlot)
library(lavaan)
library(lavaan.survey)
library(semTools)

library(ggplot2, warn.conflicts = FALSE)
library(stringr)
library(RColorBrewer)
library(MplusAutomation)
```

## Data 

Write a report on the analysis (10 pages not including
appendices & references)

• Theoretical framing of the research problem / Research questions / Hypotheses  
• Short description of the dataset used  
• Modeling strategy  
• Results  

```{r data, include=TRUE}

setwd("C:/Users/pamel/Documents/ESS/")
load("ESS5-9Round.RData")

items_o <- c("iphlppl", "iplylfr", "ipeqopt", "ipudrst", "impenv") 

vars <- c(items_o, 
          "cntry", "dweight", "hhmmb", "gndr", "agea", "GINI","GDP", "HDI",
          "dvrcdeva", "domicil", "eisced","name","essround","idno")
cont <- c("hhmmb","agea")
cat <- vars[which(!(vars%in%c(cont,"name","essround","idno","cntry","dweight")))]


```

```{r descr}

round <- c(8,9) 
countries <- c("Austria","Belgium","Czechia","Estonia","France","Germany",
               "Ireland","Italy","Netherlands","Norway","Poland","Slovenia","Switzerland","United Kingdom")
#"Hungary", "Finland"
ds_filtradaAll <- ds %>% filter(cntry %in% countries & essround %in% round) %>% 
  select(vars)
ds_filtradaAll <- copy_labels(ds_filtradaAll, ds)

table(as_character(ds_filtradaAll$cntry),ds_filtradaAll$essround)

by(ds_filtradaAll, ds_filtradaAll$essround,function(x) describe(x))

dat2 <- data.frame(reverse.code(keys = rep(-1,5), items = ds_filtradaAll[,items_o], mini = rep(1,5), maxi = rep(6,5)))
colnames(dat2) <- paste(items_o,"_r",sep = "")
labels = num_lab("
             1 Not like me at all
             2 Not like me    
             3 A little like me    
             4 Somewhat like me
             5 Like me
             6 Very much like me
")
val_lab(dat2$iphlppl_r) <- labels
val_lab(dat2$iplylfr_r) <- labels
val_lab(dat2$ipeqopt_r) <- labels
val_lab(dat2$ipudrst_r) <- labels
val_lab(dat2$impenv_r) <- labels
var_lab(dat2$iphlppl_r) <- var_lab(ds_filtradaAll$iphlppl)
var_lab(dat2$iplylfr_r) <- var_lab(ds_filtradaAll$iplylfr)
var_lab(dat2$ipeqopt_r) <- var_lab(ds_filtradaAll$ipeqopt)
var_lab(dat2$ipudrst_r) <- var_lab(ds_filtradaAll$ipudrst)
var_lab(dat2$impenv_r) <- var_lab(ds_filtradaAll$impenv)

ds_filtradaAll <- cbind(ds_filtradaAll,dat2)
items <- paste(items_o,"_r",sep = "")
for (j in round){
  for (i in items){
    print(paste(i,": ", var_lab(eval(parse(text=paste("ds_filtradaAll$",i))))))
    print(use_labels(ds_filtradaAll[ds_filtradaAll$essround == j,],
                    table(eval(parse(text=paste("ds_filtradaAll$",i))), as.character(ds_filtradaAll$cntry),useNA = "ifany")))
    print(use_labels(ds_filtradaAll[ds_filtradaAll$essround == j,],
                    round(prop.table(table(eval(parse(text=paste("ds_filtradaAll$",i))),as.character(ds_filtradaAll$cntry)),1)*100,2)))
  }
}

#Assign weight and survey stucture for ESS data
ds_filtradaAll %>% group_by(essround,cntry) %>% 
  summarise(pesos=round(sum(dweight),0), n=n(), diff=n-pesos) %>%  
  summarise(Diff_Pesos_N=sum(diff))

```

```{r recodvar}

ds_filtradaAll$gndrD <- ifelse(ds_filtradaAll$gndr == 1, 0, 
                               ifelse(ds_filtradaAll$gndr == 2, 1,ds_filtradaAll$gndr))
var_lab(ds_filtradaAll$gndrD) <- "Gender (Female)"
use_labels(ds_filtradaAll,table(gndrD,as.character(cntry)))


val_lab(ds_filtradaAll$eisced)
ds_filtradaAll$eiscedT <- ifelse(ds_filtradaAll$eisced %in% c(1,2,3) , 1,
                              ifelse(ds_filtradaAll$eisced %in% c(4,5),2,
                                     ifelse(ds_filtradaAll$eisced %in% c(6,7), 3,NA)))
val_lab(ds_filtradaAll$eiscedT) = num_lab("
            1 Less than Upper secondary
            2 Upper secondary or vocational
            3 Bachelor or higher
")
var_lab(ds_filtradaAll$eiscedT) <- var_lab(ds_filtradaAll$eisced)
use_labels(ds_filtradaAll,table(eiscedT,as.character(cntry)))
eiscedD <- as.dichotomy(ds_filtradaAll$eiscedT, prefix="eisced")
names(eiscedD)

val_lab(ds_filtradaAll$domicil)
ds_filtradaAll$domicilT <- ifelse(ds_filtradaAll$domicil %in% c(4,5) , 1,
                                  ifelse(ds_filtradaAll$domicil %in% c(3) , 2,
                                         ifelse(ds_filtradaAll$domicil %in% c(2),3,
                                                ifelse(ds_filtradaAll$domicil %in% c(1),4,NA))))

val_lab(ds_filtradaAll$domicilT) <- num_lab("
             1 Countryside 
             2 Town or small city
             3 Suburbs or outskirts of big city    
             4 A big city
")
var_lab(ds_filtradaAll$domicilT) <- var_lab(ds_filtradaAll$domicil)
use_labels(ds_filtradaAll,table(domicilT,as.character(cntry)))
domicilD <- as.dichotomy(ds_filtradaAll$domicilT, prefix="domicil")
names(domicilD)

ds_filtradaAll <- cbind(ds_filtradaAll, eiscedD, domicilD)
ds_filtradaAll <- ds_filtradaAll[,!colnames(ds_filtradaAll) %in% c("eisced55")]


ds_filtradacntry <- ds_filtradaAll %>% group_by(essround,cntry) %>% 
  summarise(n = n(),
            CntryAge = mean(agea, na.rm = TRUE),
            CntryFemale = sum(gndrD, na.rm = TRUE)/ n,
            CntryEisced1 = sum(eisced1,na.rm=TRUE)/ n,
            CntryEisced2 = sum(eisced2,na.rm=TRUE)/ n,
            CntryEisced3 = sum(eisced3,na.rm=TRUE)/ n,
            CntryDomici1 = sum(domicil1,na.rm=TRUE)/ n,
            CntryDomici2 = sum(domicil2,na.rm=TRUE)/ n,
            CntryDomici3 = sum(domicil3,na.rm=TRUE)/ n,
            CntryDomici4 = sum(domicil4,na.rm=TRUE)/ n) %>% select(-n)

ds_filtradaAll <- left_join(ds_filtradaAll,ds_filtradacntry, by=c("essround","cntry"))

ds_filtradamplus <- ds_filtradaAll
ds_filtradamplus$cntry <- as.numeric(as.factor(ds_filtradaAll$cntry))
ds_filtradamplus$essround <- as.numeric(as.factor(ds_filtradaAll$essround))
write.table(ds_filtradamplus[ds_filtradamplus$essround == 2,
                             c("cntry","dweight","HDI","iphlppl_r","iplylfr_r","ipeqopt_r","ipudrst_r","impenv_r",
                               "agea","gndrD","eisced2","eisced3","domicil2","domicil3","domicil4","CntryAge","CntryFemale")],"Sem9.dat",
            row.names = FALSE, col.names = FALSE, sep = "\t", quote = FALSE, na = ".")
write.table(ds_filtradamplus[ds_filtradamplus$essround == 1,
                             c("cntry","dweight","HDI","iphlppl_r","iplylfr_r","ipeqopt_r","ipudrst_r","impenv_r",
                               "agea","gndrD","eisced2","eisced3","domicil2","domicil3","domicil4","CntryAge","CntryFemale")],"Sem8.dat",
            row.names = FALSE, col.names = FALSE, sep = "\t", quote = FALSE, na = ".")

```

## Model CFA

```{r model, warning=FALSE}

# countries <- c("Belgium", "Germany", "Ireland", "United Kingdom")
# ds_filtradaAll <- ds_filtradaAll1 %>% filter(cntry %in% countries)

#1st order 
model13<-'
Benev =~ iphlppl_r + iplylfr_r 
Unive =~ ipeqopt_r + ipudrst_r + impenv_r 
Benev ~~ Unive
'
# model23<-'
# STrasc =~ iphlppl_r + iplylfr_r + ipeqopt_r + ipudrst_r + impenv_r 
# '

# #2nd order
# model23<-'
# Benev =~ iphlppl_r + iplylfr_r 
# Unive =~ ipeqopt_r + ipudrst_r +impenv_r
# STrasc =~ Benev + Unive
# STrasc ~~ 1*STrasc
# iplylfr_r ~~  impenv_r
# '
# 
# #Bi factor
# model33 <-'
# Benev =~ iphlppl_r + iplylfr_r 
# Unive =~ ipeqopt_r + ipudrst_r +impenv_r 
# STrasc =~ iphlppl_r + iplylfr_r  + ipeqopt_r + ipudrst_r + impenv_r
# STrasc ~~ 
# '


for (r in c(8,9)) {
  ds_filtrada <- ds_filtradaAll %>% filter(essround == r)
  survey.design <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada)
  
  lavaan.fit3 <- lavaan(model13, data=ds_filtrada, auto.fix.first=TRUE,
                       auto.var=TRUE, int.ov.free=TRUE,
                       auto.cov.lv.x=TRUE,estimator="MLM",
                       cluster = "cntry", meanstructure=TRUE)
  survey.fit3 <- lavaan.survey(lavaan.fit=lavaan.fit3,survey.design=survey.design)
  #assign(paste0("survey.fit3r",r),survey.fit3)
  
  print(paste("ESS round: ", r))
  print(fitMeasures(survey.fit3, c("chisq","pvalue","cfi", "tli","rmsea", "srmr",                                       "chisq.scaled","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust","srmr_bentler")))
  print(modindices(survey.fit3,sort=T)[1:10,])
  
  cov <- round(cov(ds_filtrada[,items], use="complete.obs"),3)
  print(lowerMat(cov, digits=3))
  print(round(colMeans(ds_filtrada[,items], na.rm = TRUE),3))
  print(fitted(survey.fit3))
  invisible(semPaths(survey.fit3,"model","std","lisrel", edge.label.cex = 0.8, intercepts = FALSE, groups = "latent", 
                     pastel = TRUE, optimizeLatRes = TRUE))
  print(summary(survey.fit3, standardized=T, rsquare=T, fit.measures=T))
}

for (r in c(8,9)) {
  ds_filtrada <- ds_filtradaAll %>% filter(essround == r)
  survey.design <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada)

  # 1. CONFIGURAL EQUIVALENCE
  ## Add the "meanstructure" argument to add means/intercepts
  lavaan.conffit3 <- lavaan(model13, data=ds_filtrada,  
                            auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                            int.ov.free=TRUE,     #intercepts not fixed to 0
                            meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                            auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                            auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                            estimator="MLM",
                            group = "cntry",
                            group.label = countries
                            #group.equal = ...    #vector for multigroup analysis specify the pattern of equality constrains across multiple groups
                            )
  survey.conffit3 <- lavaan.survey(lavaan.fit=lavaan.conffit3,survey.design=survey.design)
  # mi1 <- modindices(survey.conffit3, sort = T, free.remove = F)
  # mi1[mi1$op == "=~",]
   
  # 2. METRIC EQUIVALENCE: set the factor loadings equal across groups
  
  lavaan.metrfit3 <- lavaan(model13, data=ds_filtrada, 
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          estimator="MLM",
                          group = "cntry",
                          group.label = countries,
                          group.equal=c("loadings") #vector for multigroup analysis specify the pattern of equality constrains across multiple groups
                          )
  survey.metrfit3 <- lavaan.survey(lavaan.fit=lavaan.metrfit3,survey.design=survey.design)
  # mi2 <- modindices(survey.metrfit3, sort = T, free.remove = F)
  # loadings <- mi2[mi2$op=="=~",]
  # loadings[order(loadings$mi,decreasing=TRUE),]

  # 3. SCALAR EQUIVALENCE: set the factor loadings and the intercepts equal across groups
  
  lavaan.scalfit3 <- lavaan(model13, data=ds_filtrada, 
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          estimator="MLM",
                          group = "cntry",
                          group.label = countries,
                          group.equal=c("loadings","intercepts"),group.partial= c("iplylfr_r ~1","ipeqopt_r~1"))
  survey.scalfit3 <- lavaan.survey(lavaan.fit=lavaan.scalfit3,survey.design=survey.design)
  # mi3 <- modindices(survey.scalfit3, sort = T, free.remove = F)
  # mi3[mi3$op == "~1",]
   
  # 4. check whether factor variances are equal across groups
  lavaan.varianfit3 <- lavaan(model13, data=ds_filtrada, 
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          estimator="MLM",
                          group = "cntry",
                          group.label = countries,
                          group.equal=c("loadings","intercepts","lv.variances"),
                          group.partial= c("iplylfr_r ~1","ipeqopt_r~1")) 
  survey.varianfit3 <- lavaan.survey(lavaan.fit=lavaan.varianfit3,survey.design=survey.design)
  # mi3 <- modindices(survey.varianfit3, sort = T, free.remove = F)
  # mi3[mi3$op == "~~",]
  
  
  invar <- data.frame(round(rbind(Configural = fitMeasures(survey.conffit3, c("cfi", "rmsea","cfi.robust", "rmsea.scaled")),
                                      Metric = fitMeasures(survey.metrfit3, c("cfi", "rmsea","cfi.robust", "rmsea.scaled")),
                                      Scalar = fitMeasures(survey.scalfit3, c("cfi", "rmsea","cfi.robust", "rmsea.scaled")),
                                      Strict = fitMeasures(survey.varianfit3, c("cfi", "rmsea","cfi.robust", "rmsea.scaled"))),3))
  dif <- invar %>%
      mutate_all(funs(abs(. - lag(.))) )
  colnames(dif) <- c("delta.cfi","delta.rmsea","delta.Robcfi","delta.Scalrmsea")
  print(paste("ESS round: ", r))
  print(cbind(invar,dif))
  
  invaraut <- measurementInvariance(model = model13, 
                        data=ds_filtrada,
                        group = "cntry",group.label = countries,
                        strict = TRUE,group.partial = c("iplylfr_r ~1") )
  
  
  # countries1 <- c("Belgium", "Germany", "Ireland", "United Kingdom","Norway")
  # invaraut1 <- measurementInvariance(model = model3, 
  #                       data=ds_filtrada,
  #                       group = "cntry",group.label = countries1,
  #                       strict = TRUE,group.partial = c("iplylfr_r ~1") )
  # 
  # countries2 <- c("Czechia",  "Austria" , "Italy"  ,  "Slovenia" ,"Poland")
  # invaraut2 <- measurementInvariance(model = model3, 
  #                       data=ds_filtrada,
  #                       group = "cntry",group.label = countries2,
  #                       strict = TRUE,group.partial = c("iplylfr_r ~1") )
  # 
  # countries3 <- c("Netherlands", "France" )
  # invaraut3 <- measurementInvariance(model = model3, 
  #                       data=ds_filtrada,
  #                       group = "cntry",group.label = countries3,
  #                       strict = TRUE,group.partial = c("iplylfr_r ~1") )
  # 
  # countries4 <- c("Estonia","Switzerland")
  # invaraut4 <- measurementInvariance(model = model3, 
  #                       data=ds_filtrada,
  #                       group = "cntry",group.label = countries4,
  #                       strict = TRUE,group.partial = c("ipeqopt_r ~1") ) 
  
  results <- partialInvariance(fit = invaraut, type = "strict")
  results$results
  
  
  assign(paste0("survey.scalfit3r",r),survey.scalfit3)
  invisible(semPaths(survey.scalfit3,"model","std","lisrel", edge.label.cex = 1.2, intercepts = FALSE, layout = "tree2",
                     panelGroups = FALSE, ask  = FALSE, groups = "latent", pastel = TRUE, exoCov = TRUE, rotation = 1))

}

```

## Model SEM

```{r semmodel, warning=FALSE}

semmodel21 <-'
Benev =~ iphlppl_r + iplylfr_r
Unive =~ ipeqopt_r + ipudrst_r + impenv_r
STrasc =~ Unive + Benev
STrasc ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4 + HDI
Unive ~~ 0.2*Unive
Benev ~~ 0.2*Benev
'

 
for (r in c(8,9)) {
  ds_filtrada2 <- ds_filtradaAll %>% filter(essround == r)
  survey.design2 <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada2)

  lavaan.semfit <- lavaan(semmodel21, data=ds_filtrada2,
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          estimator="MLM",
                          cluster = "cntry")
  survey.semfit <- lavaan.survey(lavaan.fit=lavaan.semfit,survey.design=survey.design2)
  assign(paste0("survey.semfit",r),survey.semfit)

  print(paste("ESS round: ", r))
  print(fitMeasures(survey.semfit, c("chisq","pvalue","cfi", "tli","rmsea", "srmr",                                       "chisq.scaled","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust","srmr_bentler")))
  print(modindices(survey.semfit,sort=T)[1:10,])
  invisible(semPaths(survey.semfit,"model","std","lisrel", edge.label.cex = 0.8, intercepts = FALSE, layout = "tree", 
                     groups = "latent", pastel = TRUE, exoCov = FALSE, optimizeLatRes = TRUE))

  print(summary(survey.semfit, standardized=T, rsquare=T, fit.measures=T))

}

# 
# sum1 <-full_join(parameterEstimates(survey.fit3r8),
#                  parameterEstimates(survey.fit3r9), 
#                  by=c("lhs", "op", "rhs"))
# sum2 <-full_join(parameterEstimates(survey.scalfit3r8),
#                  parameterEstimates(survey.scalfit3r9), 
#                  by=c("lhs", "op", "rhs", "block", "group"))
# sum2$block <- as.character(sum2$block)
# 
# sum3 <-full_join(parameterEstimates(survey.semfit8),
#                  parameterEstimates(survey.semfit9), 
#                  by=c("lhs", "op", "rhs"))
# sum4 <-full_join(parameterEstimates(survey.semscalfit3r8),
#                   parameterEstimates(survey.semscalfit3r9), 
#                   by=c("lhs", "op", "rhs", "block", "group"))
# sum2 <- sum2 %>%  mutate(est.x = ifelse(pvalue.x > 0.05, NA, round(est.x,3)),
#                          est.y = ifelse(pvalue.y > 0.05, NA, round(est.y,3)))
# 
# val_lab(sum2$block) <- cntrylabels
# sum2$block <- as.character(sum2$block)
# 
# sum4 <- sum4 %>%  mutate(est.x = ifelse(pvalue.x > 0.05, NA, round(est.x,3)),
#                          est.x = ifelse(rhs == "agea", est.x*10, est.x),
#                          est.y = ifelse(pvalue.y > 0.05, NA, round(est.y,3)),
#                          est.y = ifelse(rhs == "agea", est.y*10, est.y),
#                          rhs1 = ifelse(rhs == "gndrD", "Gender (Female / Male)",
#                                 ifelse(rhs == "agea", "Age (10 years increment)",
#                                 ifelse(rhs == "eisced2", "Highest level of education, (Upper secondary or vocational / Less than Upper secondary)", 
#                                 ifelse(rhs == "eisced3", "Highest level of education, (Bachelor or higher / Less than Upper secondary)", 
#                                 ifelse(rhs == "domicil2", "Domicile (Town or small city/Countryside)", 
#                                 ifelse(rhs == "domicil3", "Domicile (Suburbs or outskirts of big city/Countryside)",
#                                 ifelse(rhs == "domicil4", "Domicile (A big city/Countryside)",
#                                 ifelse(rhs == "HDI_i", "Human Development Index", rhs)))))))))
# val_lab(sum4$block) <- cntrylabels
# sum4$block <- as.character(sum4$block)
# 
# dir <- "G:/My Drive/Master in Statistics/Structural equations/Paper/"
# write.table(sum1,paste0(dir,"Parametersfit.csv"), sep = ",", row.names = FALSE)
# write.table(sum2,paste0(dir,"ParametersScalfit.csv"), sep = ",", row.names = FALSE)
# 
# write.table(sum3,paste0(dir,"ParametersSemfit.csv"), sep = ",", row.names = FALSE)
# write.table(sum4,paste0(dir,"ParametersSemScalfit.csv"), sep = ",", row.names = FALSE)
```

## Ordered variables

The model with categorical variables is undefined, only 5 variables and too many parameters to estimate.

```{r Ordered, eval=FALSE}

for (r in c(8,9)) {
  ds_filtrada <- ds_filtradaAll %>% filter(essround == r)
  survey.design <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada)
  
  lavaan.Ordfit3 <- lavaan(model13, data=ds_filtrada, estimator = "WLSMV",
                       cluster = "cntry", 
                      ordered = c("iphlppl_r", "iplylfr_r", "ipeqopt_r", "ipudrst_r", "impenv_r"))
  survey.Ordfit3 <- lavaan.Ordfit3
  #survey.Ordfit3 <- lavaan.survey(lavaan.fit=lavaan.Ordfit3,survey.design=survey.design)
  assign(paste0("survey.Ordfit3r",r),survey.Ordfit3)
  
  print(paste("ESS round: ", r))
  print(fitMeasures(survey.Ordfit3, c("chisq","pvalue","cfi", "tli","rmsea", "srmr",
                                      "chisq.scaled","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust","srmr_bentler")))
  #print(modindices(survey.Ordfit3,sort=T)[1:10,])
  
  # cov <- round(cov(ds_filtrada[,items], use="complete.obs"),3)
  # print(lowerMat(cov, digits=3))
  # print(round(colMeans(ds_filtrada[,items], na.rm = TRUE),3))
  # print(fitted(survey.Ordfit3))
  # invisible(semPaths(survey.Ordfit3,"model","stand", style = "lisrel", rainbowStart = 0.8))
  # 
  print(summary(survey.Ordfit3, standardized=T, rsquare=T, fit.measures=T))
}

for (r in c(8,9)) {
  ds_filtrada <- ds_filtradaAll %>% filter(essround == r)
  survey.design <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada)
  
  lavaan.Ordfit3 <- lavaan(model13, data=ds_filtrada, estimator = "WLSMV",
                        ordered = c("iphlppl_r", "iplylfr_r", "ipeqopt_r", "ipudrst_r", "impenv_r"))
  survey.Ordfit3 <- lavaan.Ordfit3
  assign(paste0("survey.Ordfit3r",r),survey.Ordfit3)
  
  print(paste("ESS round: ", r))
  print(fitMeasures(survey.Ordfit3, c("chisq","pvalue","cfi", "tli","rmsea", "srmr",
                                      "chisq.scaled","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust","srmr_bentler")))
  #print(modindices(survey.Ordfit3,sort=T)[1:10,])
  
  # cov <- round(cov(ds_filtrada[,items], use="complete.obs"),3)
  # print(lowerMat(cov, digits=3))
  # print(round(colMeans(ds_filtrada[,items], na.rm = TRUE),3))
  # print(fitted(survey.Ordfit3))
  # invisible(semPaths(survey.Ordfit3,"model","stand", style = "lisrel", rainbowStart = 0.8))
  # 
  print(summary(survey.Ordfit3, standardized=T, rsquare=T, fit.measures=T))
}
```

## Multilevel CFA

```{r mmul, warning=FALSE}

mmodel31<-'
level: 1
Benev_w =~ iphlppl_r + iplylfr_r
Unive_w =~ ipeqopt_r + ipudrst_r + impenv_r
Unive_w ~~ Benev_w
Unive_w ~~ 0.3*Unive_w
Benev_w ~~ 0.4*Benev_w
level: 2
Benev_b =~ iphlppl_r + iplylfr_r
Unive_b =~ ipeqopt_r + ipudrst_r
Unive_b ~~ Benev_b
Unive_b ~~ 0.03*Unive_b
Benev_b ~~ 0.04*Benev_b
'


#+ impenv_r
for (r in c(8,9)) {
  ds_filtrada <- ds_filtradaAll %>% filter(essround == r)
  
  lavaan.mfit3 <- lavaan(mmodel31, data=ds_filtrada, auto.fix.first=TRUE, sampling.weights = "dweight",
                       auto.var=TRUE, int.ov.free=TRUE, auto.cov.lv.x=TRUE, 
                       cluster = "cntry", meanstructure=TRUE)
  survey.mfit3 <- lavaan.mfit3
  assign(paste0("survey.mfit3r",r),survey.mfit3)
  
  print(paste("ESS round: ", r))

  print(fitMeasures(survey.mfit3, c("chisq","pvalue","cfi", "tli","rmsea", "srmr")))
  print(modindices(survey.mfit3,sort=T)[1:10,])
  
  cov <- round(cov(ds_filtrada[,items], use="complete.obs"),3)
  print(lowerMat(cov, digits=3))
  print(round(colMeans(ds_filtrada[,items], na.rm = TRUE),3))
  print(fitted(survey.mfit3))
  
  print(summary(survey.mfit3, standardized=T, rsquare=T, fit.measures=T))

 invisible(semPaths(paste0("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\mcfa",r,".out"), "model", "std", nCharNodes = 0,mar = c(5, 1, 5, 1),
           style = "lisrel",  edge.label.cex = 0.8, intercepts = FALSE, layout = "tree", 
                     groups = "latent", pastel = TRUE, exoCov = FALSE, optimizeLatRes = TRUE, ask = FALSE))

}
```

# Multilevel SEM

```{r Msem}

# Msemmodel <-'
# level: 1 
# STrasc_w =~ iphlppl_r + iplylfr_r + ipeqopt_r + ipudrst_r + impenv_r 
# STrasc_w ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4
# level: 2
# STrasc_b =~ iphlppl_r + iplylfr_r + ipeqopt_r + ipudrst_r + impenv_r 
# STrasc_b ~  HDI
# '
Msemmodel<-'
level: 1
Benev_w =~ iphlppl_r + iplylfr_r
Unive_w =~ ipeqopt_r + ipudrst_r + impenv_r
STrasc_w =~ 1*Unive_w + 1*Benev_w
STrasc_w ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4
STrasc_w ~~ 0.3*STrasc_w
Benev_w ~~ 0.2*Benev_w
Unive_w ~~ 0.2*Unive_w
level: 2
Benev_b =~ iphlppl_r + iplylfr_r
Unive_b =~ ipeqopt_r + ipudrst_r + impenv_r
STrasc_b =~ Unive_b + Benev_b
STrasc_b ~  HDI
iphlppl_r ~ 0 
impenv_r ~ 0 
ipudrst_r ~ 0 
iplylfr_r ~ 0 
ipeqopt_r ~ 0
STrasc_b ~~ 0.04*STrasc_b 
Benev_b ~~ 0.03*Benev_b 
Unive_b ~~ 0.03*Unive_b
'


for (r in c(8,9)) {
  ds_filtrada2 <- ds_filtradaAll %>% filter(essround == r)
   
  lavaan.Msemfit <- lavaan(Msemmodel, data=ds_filtrada2, sampling.weights = "dweight",
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          cluster = "cntry")
  assign(paste0("survey.Msemfit",r),lavaan.Msemfit)

  print(paste("ESS round: ", r))
  print(fitMeasures(lavaan.Msemfit, c("chisq","pvalue","cfi", "tli","rmsea", "srmr")))
  print(modindices(lavaan.Msemfit,sort=T)[1:10,])
  
  print(summary(lavaan.Msemfit, standardized=T, rsquare=T, fit.measures=T))
  invisible(semPaths(paste0("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\msem",r,".out"), "model", "std", nCharNodes = 0,mar = c(5, 1, 5, 1),
           style = "lisrel",  edge.label.cex = 0.8, intercepts = FALSE, layout = "tree", 
                     groups = "latent", pastel = TRUE, exoCov = FALSE, optimizeLatRes = TRUE, ask = FALSE))


  
}


# 
# 
# cntrylabels <- num_lab("
#   1	Austria
#   2	Belgium
#   3	Czechia
#   4	Estonia
#   5	France
#   6	Germany
#   7	Ireland
#   8	Italy
#   9	Netherlands
#   10	Norway
#   11	Poland
#   12	Slovenia
#   13	Switzerland
#   14	United Kingdom"
# )
# 
sum1 <-full_join(parameterEstimates(survey.semfit),
                 parameterEstimates(survey.semfit),
                 by=c("lhs", "op", "rhs"))
sum1 <- sum1 %>%  mutate(est.x = ifelse(pvalue.x > 0.05, NA, round(est.x,3)),
                         est.x = ifelse(rhs == "agea", est.x*10, est.x),
                         est.y = ifelse(pvalue.y > 0.05, NA, round(est.y,3)),
                         est.y = ifelse(rhs == "agea", est.y*10, est.y),
                         rhs1 = ifelse(rhs == "gndrD", "Gender (Female / Male)",
                                ifelse(rhs == "agea", "Age (10 years increment)",
                                ifelse(rhs == "eisced2", "Highest level of education, (Upper secondary or vocational / Less than Upper secondary)",
                                ifelse(rhs == "eisced3", "Highest level of education, (Bachelor or higher / Less than Upper secondary)",
                                ifelse(rhs == "domicil2", "Domicile (Town or small city/Countryside)",
                                ifelse(rhs == "domicil3", "Domicile (Suburbs or outskirts of big city/Countryside)",
                                ifelse(rhs == "domicil4", "Domicile (A big city/Countryside)",
                                ifelse(rhs == "HDI","Human Development Index",rhs)))))))))
sum3 <-full_join(parameterEstimates(survey.Msemfit8),
                 parameterEstimates(survey.Msemfit9),
                 by=c("lhs", "op", "rhs","block","level"))
sum3 <- sum3 %>%  mutate(est.x = ifelse(pvalue.x > 0.05, NA, round(est.x,3)),
                         est.x = ifelse(rhs == "agea", est.x*10, est.x),
                         est.y = ifelse(pvalue.y > 0.05, NA, round(est.y,3)),
                         est.y = ifelse(rhs == "agea", est.y*10, est.y),
                         rhs1 = ifelse(rhs == "gndrD", "Gender (Female / Male)",
                                ifelse(rhs == "agea", "Age (10 years increment)",
                                ifelse(rhs == "eisced2", "Highest level of education, (Upper secondary or vocational / Less than Upper secondary)",
                                ifelse(rhs == "eisced3", "Highest level of education, (Bachelor or higher / Less than Upper secondary)",
                                ifelse(rhs == "domicil2", "Domicile (Town or small city/Countryside)",
                                ifelse(rhs == "domicil3", "Domicile (Suburbs or outskirts of big city/Countryside)",
                                ifelse(rhs == "domicil4", "Domicile (A big city/Countryside)",
                                ifelse(rhs == "HDI","Human Development Index",rhs)))))))))

dir <- "G:/My Drive/Master in Statistics/Structural equations/Paper/"
write.table(sum1,paste0(dir,"ParametersCSemfit.csv"), sep = ",", row.names = FALSE)
write.table(sum3,paste0(dir,"ParametersMSemfit.csv"), sep = ",", row.names = FALSE)
```

# MIMIC 

```{r mimic, warning=FALSE}

mimicmodel21 <-'
Benev =~ iphlppl_r + iplylfr_r
Unive =~ ipeqopt_r + ipudrst_r + impenv_r
Benev ~~ Unive
Benev ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4 + HDI
Unive ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4 + HDI
'

 
for (r in c(8,9)) {
  ds_filtrada2 <- ds_filtradaAll %>% filter(essround == r)
  survey.design2 <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada2)

  lavaan.mimicfit <- lavaan(mimicmodel21, data=ds_filtrada2,
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          estimator="MLM",
                          cluster = "cntry")
  survey.mimicfit <- lavaan.survey(lavaan.fit=lavaan.mimicfit,survey.design=survey.design2)
  assign(paste0("survey.mimicfit",r),survey.mimicfit)

  print(paste("ESS round: ", r))
  print(fitMeasures(survey.mimicfit, c("chisq","pvalue","cfi", "tli","rmsea", "srmr",                                       "chisq.scaled","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust","srmr_bentler")))
  print(modindices(survey.mimicfit,sort=T)[1:10,])
  invisible(semPaths(survey.mimicfit,"model","std","lisrel", edge.label.cex = 0.8, intercepts = FALSE, layout = "tree", 
                     groups = "latent", pastel = TRUE, exoCov = FALSE, optimizeLatRes = TRUE, rotation = 2))

  print(summary(survey.mimicfit, standardized=T, rsquare=T, fit.measures=T))

}
```


# Multilevel MIMIC 

```{r mmimic, warning=FALSE}

Mmimicmodel21 <-'
level: 1
Benev_w =~ iphlppl_r + iplylfr_r
Unive_w =~ ipeqopt_r + ipudrst_r + impenv_r
Benev_w ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4
Unive_w ~  agea + gndrD + eisced2 + eisced3 + domicil2 + domicil3 + domicil4
level: 2
Benev_b =~ iphlppl_r + iplylfr_r
Unive_b =~ ipeqopt_r + ipudrst_r + impenv_r
Benev_b ~  HDI
Unive_b ~  HDI
'
 
for (r in c(8,9)) {
  ds_filtrada2 <- ds_filtradaAll %>% filter(essround == r)
  survey.design2 <- svydesign(ids=~idno, prob=~dweight, data=ds_filtrada2)

  lavaan.Mmimicfit <- lavaan(Mmimicmodel21, data=ds_filtrada2,sampling.weights = "dweight",
                          auto.fix.first=TRUE,  #factor loading of first indicator set to 1
                          int.ov.free=TRUE,     #intercepts not fixed to 0
                          meanstructure=TRUE,   #the means of the observed variables enter the model, not what user define
                          auto.var=TRUE,        #residual variances and variances of exogeneous latent variables are included in the model and set free
                          auto.cov.lv.x=TRUE,   #covariances of exogeneous latent variables are included in the model and set free
                          estimator="MLM",
                          cluster = "cntry")
  survey.Mmimicfit <- lavaan.Mmimicfit
  assign(paste0("survey.Mmimicfit",r),survey.Mmimicfit)

  print(paste("ESS round: ", r))
  print(fitMeasures(survey.Mmimicfit, c("chisq","pvalue","cfi", "tli","rmsea", "srmr",                                       "chisq.scaled","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust","srmr_bentler")))
  print(modindices(survey.Mmimicfit,sort=T)[1:10,])
  invisible(semPaths(paste0("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\mmimic",r,".out"),"model","std","lisrel", edge.label.cex = 0.8, 
                     intercepts = FALSE, layout = "tree2", ask =FALSE,
                     groups = "latent", pastel = TRUE, exoCov = FALSE, optimizeLatRes = TRUE, rotation = 2))

  print(summary(survey.Mmimicfit, standardized=T, rsquare=T, fit.measures=T))

}
```

## Results

```{r results, warning=FALSE, echo= FALSE}
a1 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\sem8.out")
a1 <-a1$parameters$stdyx.standardized
a1 <- plyr::rename(a1,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))
a2 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\sem9.out")
a2 <-a2$parameters$stdyx.standardized
a2 <- plyr::rename(a2,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))
b1 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\msem9.out")
b1 <-b1$parameters$std.standardized
b1 <- plyr::rename(b1,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))
b2 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\msem9.out")
b2 <-b2$parameters$std.standardized
b2 <- plyr::rename(b2,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))

coef1<-rbind(cbind(ESS="ESS8",a1),
            cbind(ESS="ESS9",a2))
coef3<-rbind(cbind(ESS="ESS8",b1),
            cbind(ESS="ESS9",b2))
coef <- rbind(cbind(Type = "SEM",BetweenWithin = "Within", coef1),cbind(Type="MLSEM",coef3))

coef <- coef %>% 
  mutate(est = ifelse(rhs == "agea", est*10, est),
         est = ifelse(pvalue > 0.05, NA, round(est,2)),
         op = ifelse(lhs %in% c("STRASC.ON","STRASC_W.ON","STRASC_B.ON"), "~", lhs),
         lhs = ifelse(lhs %in% c("STRASC.ON","STRASC_W.ON","STRASC_B.ON"), "Self-Transcendence", lhs),
         rhs1 = ifelse(tolower(rhs) == "gndrd", "Gender (Female / Male)",
                ifelse(tolower(rhs) == "agea", "Age (10 years increment)",
                ifelse(tolower(rhs) == "eisced2", "Highest level of education, (Upper secondary or vocational / Less than Upper secondary)", 
                ifelse(tolower(rhs) == "eisced3", "Highest level of education, (Bachelor or higher / Less than Upper secondary)", 
                ifelse(tolower(rhs) == "domicil2", "Domicile (Town or small city/Countryside)", 
                ifelse(tolower(rhs) == "domicil3", "Domicile (Suburbs or outskirts of big city/Countryside)",
                ifelse(tolower(rhs) == "domicil4", "Domicile (A big city/Countryside)",
                ifelse(rhs == "HDI","Human Development Index",rhs)))))))),
         rhs1 = stringr::str_wrap(rhs1,30),
         order = ifelse(tolower(rhs) == "gndrD",8, ifelse( tolower(rhs) == "agea", 7, ifelse(tolower(rhs) == "eisced2", 6, ifelse(tolower(rhs) == "eisced3", 5, 
                 ifelse(tolower(rhs) == "domicil2", 4, ifelse(tolower(rhs) == "domicil3", 3, ifelse(tolower(rhs) == "domicil4", 2, ifelse(rhs == "HDI", 1,NA))))))))) %>% 
  filter(op == "~")

ggplot(coef,aes(x=reorder(rhs1,order), y=est, group = Type, fill=Type, label = est)) +
  geom_bar( stat="identity", position=position_dodge()) + 
  coord_flip() +
  facet_grid(lhs~ESS) +
  scale_y_continuous(breaks = c(-0.4,-0.2,0,0.2,0.4,0.6,0.8,1), limits = c(-0.4,1)) +
  geom_hline(yintercept=0, color = "black", size=1) +
  geom_text(aes(label = est,hjust = ifelse(coef$est > 0, 0, 1)), position = position_dodge(width = 1),  color ="black", size=3) +
  theme(legend.position = "top") +
  xlab("Sociodemographic characteristics") + ylab("Estimates coefficient") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Coefficient regression")

a1 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\mimic8.out")
a1 <-a1$parameters$std.standardized
a1 <- plyr::rename(a1,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))
a2 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\mimic9.out")
a2 <-a2$parameters$std.standardized
a2 <- plyr::rename(a2,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))
b1 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\mmimic9.out")
b1 <-b1$parameters$std.standardized
b1 <- plyr::rename(b1,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))
b2 <- readModels("C:\\Users\\pamel\\Documents\\ESS\\MPLUS\\mmimic9.out")
b2 <-b2$parameters$std.standardized
b2 <- plyr::rename(b2,c("param" = "rhs","paramHeader"="lhs", "pval"="pvalue"))


coef2<-rbind(cbind(ESS="ESS8", a1),
             cbind(ESS="ESS9", a2))
coef4<-rbind(cbind(ESS="ESS8",b1),
            cbind(ESS="ESS9",b2))
coeff <- rbind(cbind(Type = "MIMIC",BetweenWithin = "Within",coef2),cbind(Type="MLMIMIC",coef4))

coeff <- coeff %>% 
  mutate(est = ifelse(rhs == "agea", est*10, est),
         est = ifelse(pvalue > 0.05, NA, round(est,2)),
         op = ifelse(lhs %in% c("BENEV.ON","BENEV_W.ON","BENEV_B.ON","UNIVE.ON","UNIVE_W.ON","UNIVE_B.ON"), "~", lhs),
         lhs = ifelse(lhs %in% c("BENEV.ON","BENEV_W.ON","BENEV_B.ON"), "Benevolence", 
                      ifelse(lhs %in% c("UNIVE.ON","UNIVE_W.ON","UNIVE_B.ON"), "Universalism", lhs)),
         rhs1 = ifelse(tolower(rhs) == "gndrd", "Gender (Female / Male)",
                ifelse(tolower(rhs) == "agea", "Age (10 years increment)",
                ifelse(tolower(rhs) == "eisced2", "Highest level of education, (Upper secondary or vocational / Less than Upper secondary)", 
                ifelse(tolower(rhs) == "eisced3", "Highest level of education, (Bachelor or higher / Less than Upper secondary)", 
                ifelse(tolower(rhs) == "domicil2", "Domicile (Town or small city/Countryside)", 
                ifelse(tolower(rhs) == "domicil3", "Domicile (Suburbs or outskirts of big city/Countryside)",
                ifelse(tolower(rhs) == "domicil4", "Domicile (A big city/Countryside)",
                ifelse(rhs == "HDI","Human Development Index",rhs)))))))),
         rhs1 = stringr::str_wrap(rhs1,30),
         order = ifelse(tolower(rhs) == "gndrD",8, ifelse( tolower(rhs) == "agea", 7, ifelse(tolower(rhs) == "eisced2", 6, ifelse(tolower(rhs) == "eisced3", 5, 
                 ifelse(tolower(rhs) == "domicil2", 4, ifelse(tolower(rhs) == "domicil3", 3, ifelse(tolower(rhs) == "domicil4", 2, ifelse(rhs == "HDI", 1,NA))))))))) %>% 
  filter(op == "~")

ggplot(coeff,aes(x=reorder(rhs1,order), y=est, group = Type, fill=Type)) +
  geom_bar( stat="identity", position=position_dodge()) + 
  coord_flip() +
  facet_grid(lhs~ESS) +
  scale_y_continuous(breaks = c(-0.4,-0.2,0,0.2,0.4,0.6,0.8,1), limits = c(-0.4,1)) +
  geom_hline(yintercept=0, color = "black", size=1) +
  geom_text(aes(label = est,hjust = ifelse(coeff$est > 0, 0, 1)), position = position_dodge(width = 1),  color ="black", size=3) +
  theme(legend.position = "top") +
  xlab("Sociodemographic characteristics") + ylab("Estimates coefficient") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Coefficient regression")

# 
# 
# nb.cols <- 14
# mycolors <- colorRampPalette(brewer.pal(4, "Set2"))(nb.cols)
# 
# print("Continuous exogeneous variables")
# coef<-rbind(cbind(ESS="ESS8",parameterEstimates(survey.semconffit3r8)),
#             cbind(ESS="ESS9",parameterEstimates(survey.semconffit3r9)))
# 
# coeffilter <- coef %>% 
#   mutate(est = ifelse(pvalue > 0.05, NA, round(est,3)),
#          est = ifelse(rhs == "agea", est*10, est),
#          rhs1 = ifelse(rhs == "gndrD", "Gender (Female / Male)",
#                 ifelse(rhs == "agea", "Age (10 years increment)",
#                 ifelse(rhs == "eisced2", "Highest level of education, (Upper secondary or vocational / Less than Upper secondary)", 
#                 ifelse(rhs == "eisced3", "Highest level of education, (Bachelor or higher / Less than Upper secondary)", 
#                 ifelse(rhs == "domicil2", "Domicile (Town or small city/Countryside)", 
#                 ifelse(rhs == "domicil3", "Domicile (Suburbs or outskirts of big city/Countryside)",
#                 ifelse(rhs == "domicil4", "Domicile (A big city/Countryside)",
#                 ifelse(rhs == "HDI","Human Development Index",rhs)))))))),
#          rhs1 = stringr::str_wrap(rhs1,30),
#          order = ifelse(rhs == "gndrD",7, ifelse( rhs == "agea", 6, ifelse(rhs == "eisced2", 5, ifelse(rhs == "eisced3", 4, 
#                  ifelse(rhs == "domicil2", 3, ifelse(rhs == "domicil3", 2, ifelse(rhs == "domicil4", 1, ifelse(rhs == "HDI", 0,NA))))))))) %>% 
#   filter(op == "~")
# 
# coeffilter1 <- coeffilter %>% filter(rhs %in% c("gndrD","agea")) 
# val_lab(coeffilter1$block) <- cntrylabels
# use_labels(coeffilter1, { 
#   ggplot(coeffilter1,aes(x=factor(block), y=est, label = est, fill = factor(block))) + 
#   geom_bar(stat="identity", show.legend = FALSE) + 
#   coord_flip() +
#   geom_text(hjust = ifelse(est >= 0, 1.02, 0.02), color ="white", size=2.5, ) +
#   facet_grid(rhs1~ESS) +
#     scale_x_discrete(limits = rev(levels(factor(block))))+
#     xlab("Countries") + ylab("Estimates coefficient") +
#     ggtitle("Personal background") +
#     scale_fill_manual(values = mycolors)
# })
# 
# coeffilter2 <- coeffilter %>% filter(str_detect(rhs,"eisced"))
# val_lab(coeffilter2$block) <- cntrylabels
# use_labels(coeffilter2, { 
#   ggplot(coeffilter2,aes(x=factor(block), y=est, label = est, fill = factor(block))) + 
#   geom_bar(stat="identity", show.legend = FALSE) + 
#   coord_flip() +
#   geom_text(hjust = ifelse(est >= 0, 1.02, 0.02), color ="white", size=2.5) +
#   facet_grid(factor(rhs1)~ESS) +
#     scale_x_discrete(limits = rev(levels(factor(block))))+
#     xlab("Countries") + ylab("Estimates coefficient")+
#     ggtitle("Educational background") +
#     scale_fill_manual(values = mycolors)
# })
# 
# coeffilter3 <- coeffilter %>% filter(str_detect(rhs,"domicil"))
# val_lab(coeffilter3$block) <- cntrylabels
# use_labels(coeffilter3, { 
#   ggplot(coeffilter3,aes(x=factor(block), y=est, label = est, fill = factor(block))) + 
#   geom_bar(stat="identity", show.legend = FALSE) + 
#   coord_flip() +
#   geom_text(hjust = ifelse(est >= 0, 1.02, 0.02), color ="white", size=2.5) +
#   facet_grid(factor(rhs1)~ESS, labeller = label_context) +
#     scale_x_discrete(limits = rev(levels(factor(block))))+
#     xlab("Countries") + ylab("Estimates coefficient") +
#     ggtitle("Geographical background") +
#     scale_fill_manual(values = mycolors)
# })

# coeffilter4 <- coeffilter %>% filter(str_detect(rhs,"HDI"))
# val_lab(coeffilter4$block) <- cntrylabels
# use_labels(coeffilter4, { 
#   ggplot(coeffilter4,aes(x=factor(block), y=est, label = est, fill = factor(block))) + 
#   geom_bar(stat="identity", show.legend = FALSE) + 
#   coord_flip() +
#   geom_text(hjust = ifelse(est >= 0, 1.02, 0.02), color ="white", size=2.5) +
#   facet_grid(.~ESS, labeller = label_context) +
#     scale_x_discrete(limits = rev(levels(factor(block))))+
#     xlab("Countries") + ylab("Estimates coefficient") +
#     ggtitle("Country characteristic") +
#     scale_fill_manual(values = mycolors)
# })

```
